---
title: "Comparison PPMI ODPC multiple methods add ICA "
author: "Cynthia Sandor"
date: 'Last update: `r date()`'
output:
  html_document:
    fig_height: 5
    fig_width: 5
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 5
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
setwd('/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_multiple_approaches')
library(phenix)
library(ggplot2)
library(gplots)
library(reshape)


load("/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/PPMI.Rdata")
load("/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/ODPC.Rdata")
load('/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/cat_odpc.Rdata')
load('/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/cat_ppmi2.Rdata')
load('/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/PPMI_Without_Genetics.Rdata')
load('/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/OPDC_Without_Genetics.Rdata')
Phen<-read.table('/Users/dpag0499/Documents/PROJECTS/PPMI/data/phenix_input/Caseonly_gemma2_pheno.txt',header=TRUE,sep='\t',row.names = 1)
cat_ppmi_red<-cat_ppmi[which(cat_ppmi$Clinical %in% colnames(Y_temp2)),]
```

```{r functions, include=FALSE}

make_plot_bar<-function(id,df_mm_s) {
  id<-which(df_mm_s$Underly_Variable==paste("V",i,sep=''))
  p<-ggplot(df_mm_s[id,], aes(x=Clinical, y=Value,fill=Cat)) + 
  geom_bar(stat='identity',position=position_dodge()) + 
    geom_bar(stat='identity',position=position_dodge()) +  
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
p
}

correlation_lat_var<-function(Y,S,n_axes,cat) {
  n_clinical<-length(colnames(Y))
  list_axes<-c(1:n_axes)
  mm <- matrix(NA,n_clinical,n_axes)

  for (i in 1:n_axes) {
    for (j in 1:n_clinical) {
      t<-cor.test(Y[,j],S[,i])
      mm[j,i]<-as.numeric(t$estimate)
    }
  }

    df_mm<-as.data.frame(mm)
    colnames(df_mm)<-paste0("V",c(1:n_axes))
    row.names(df_mm)<-colnames(Y)
    df_mm$Clinical<-row.names(df_mm)
    df_mm$Cat<-cat$Cat
    df_mm_s<-melt(df_mm,id=c("Clinical","Cat"))
    colnames(df_mm_s)[4]<-"Value"
    colnames(df_mm_s)[3]<-"Underly_Variable"
    df_mm_s$Clinical <- factor(df_mm_s$Clinical, levels = unique(df_mm_s$Clinical[order(df_mm_s$Cat)]))
    df_mm_s
    
}



correlation_table<-function(df_phenix,cat_ppmi) {

col_cat=c()
col_S=c()
col_x=c()
col_y=c()
n_cat<-length(unique(cat_ppmi$Cat))
list_cat<-unique(cat_ppmi$Cat)
  for (i in 1:3) {
    for (j in 1:n_cat) {
     S<-paste("V",i,sep="")
      index<-which(df_phenix$Underly_Variable==S & df_phenix$Cat==list_cat[j] & df_phenix$cohort=="ODPC")
      a<-mean(df_phenix$Value[index])
      index<-which(df_phenix$Underly_Variable==S & df_phenix$Cat==list_cat[j]  & df_phenix$cohort=="PPMI")
      b<-mean(df_phenix$Value[index])
      col_S<-c(S,col_S)
      col_cat<-c(col_cat,list_cat[j])
      col_x<-c(col_x,a)
      col_y<-c(col_y,b)
    }
  }
  df_phenix_comp<-data.frame(S=col_S,Cat=col_cat,ODPC=col_x,PPMI=col_y)
 }
```


## Compute MDS/PCA/ICA/phenix

I performed MDS/PCA/ICA analyses on imputed phenotype (I can replicate the result below by replacing missing phenotype by mean but only the phenotype with < 5% have been considered)
For I performed MDS with two dimensions and I considered in further analyses two first PCA and 2 first latent variable of phenix.

### ICA
```{r ICA ,echo=FALSE}
library(fastICA)
ica<-fastICA(out$imp, n.comp=3)
ica_opdc<-ica$S

ica<-fastICA(out2$imp, n.comp=3)
ica_ppmi<-ica$S
```

### MDS
```{r MDS ,echo=FALSE}
d <- dist(out$imp) # euclidean distances between the rows
test<- cmdscale(d,eig=TRUE,k=3)
mds_odpc <-test$points

d <- dist(out2$imp) # euclidean distances between the rows
test<-cmdscale(d,eig=TRUE,k=3)
mds_ppmi <-test$points
mds_ppmi[,1]<--mds_ppmi[,1]
```

### PCA
```{r PCA ,echo=FALSE}
test<-prcomp(out$imp,scale=T)
pca_odpc<-test$x
test<-prcomp(out2$imp,scale=T)
pca_ppmi<-test$x
pca_ppmi[,1]<-pca_ppmi[,1]

```

### Phenix
```{r phenix 0 ,echo=FALSE}
phenix0_odpc<-out_without_genetics$S
phenix0_ppmi<-out_without_genetics_ppmi$S
phenix0_ppmi[,2]<--phenix0_ppmi[,2]
phenix0_ppmi[,3]<-phenix0_ppmi[,4]
```

### Phenix 
```{r phenix ,echo=FALSE}
phenix_odpc<-out$S
phenix_ppmi<-out2$S
phenix_ppmi[,2]<--phenix_ppmi[,2]
phenix_ppmi[,3]<-phenix_ppmi[,4]
```


## Compute correlation underlying variable & phenotypic axes

As PPMI and ODPC do not use the same clinical assessments, I classified into different categories:
(1) Motors
(2) Cognitive
(3) Anxiety
(4) Autonomic
(5) Sleep
(6) Others

These categories have been defined by PPMI cnsortium.

(1) I computed the correlation (pearson) between clinical assements with principal variable for different categories of phenotype in both cohort. 
(2) I computed the mean of correlation by category of phenotype
(3) I compared these correlation by category p




###  Phenix
```{r Phenix correlation ,echo=FALSE}
df_phenix_odpc<-correlation_lat_var(out$imp,phenix_odpc,3,cat_opdc)
df_phenix_ppmi<-correlation_lat_var(out2$imp,phenix_ppmi,3,cat_ppmi_red)

df_phenix<-rbind(df_phenix_odpc,df_phenix_ppmi)
df_phenix$cohort<-c(rep("ODPC",length(row.names(df_phenix_odpc))),rep("PPMI",length(row.names(df_phenix_ppmi))))
df_phenix$S<-paste0(df_phenix$Underly_Variable,sep="_",df_phenix$Cat)
df_phenix_mean<-correlation_table(df_phenix,cat_ppmi_red)
colnames(df_phenix_mean)[3]<-"OPDC"
p <- ggplot(df_phenix_mean, aes(OPDC, PPMI))+ geom_point(aes(colour = factor(S)),size=10
                                                         ) +geom_text(aes(label=Cat),size=7,hjust=0, vjust=0) + xlim(-0.6, 0.6)+ ylim(-0.6,0.6)+ geom_abline(intercept =0, slope = 1)+theme(axis.text=element_text(size=10),axis.title=element_text(size=18))+ theme(legend.position="none")
p
p_phenix<-p
cor.test(df_phenix_mean$OPDC,df_phenix_mean$PPMI)

```

###  Phenix 0
```{r Phenix correlation ,echo=FALSE}
df_phenix0_odpc<-correlation_lat_var(out$imp,phenix0_odpc,3,cat_opdc)
df_phenix0_ppmi<-correlation_lat_var(out2$imp,phenix0_ppmi,3,cat_ppmi_red)

df_phenix0<-rbind(df_phenix0_odpc,df_phenix0_ppmi)
df_phenix0$cohort<-c(rep("ODPC",length(row.names(df_phenix0_odpc))),rep("PPMI",length(row.names(df_phenix0_ppmi))))
df_phenix0$S<-paste0(df_phenix0$Underly_Variable,sep="_",df_phenix0$Cat)
df_phenix0_mean<-correlation_table(df_phenix0,cat_ppmi_red)
colnames(df_phenix0_mean)[3]<-"OPDC"
p <- ggplot(df_phenix0_mean, aes(OPDC, PPMI))+ geom_point(aes(colour = factor(S)),size=10
                                                         ) +geom_text(aes(label=Cat),size=7,hjust=0, vjust=0) + xlim(-0.6, 0.6)+ ylim(-0.6,0.6)+ geom_abline(intercept =0, slope = 1)+theme(axis.text=element_text(size=10),axis.title=element_text(size=18))+ theme(legend.position="none")
p
p_phenix0<-p
cor.test(df_phenix0_mean$OPDC,df_phenix0_mean$PPMI)

```

###  ICA
```{r ICA correlation ,echo=FALSE}
df_ica_opdc<-correlation_lat_var(out$imp,ica_opdc,3,cat_opdc)
df_ica_ppmi<-correlation_lat_var(out2$imp,ica_ppmi,3,cat_ppmi_red)

df_ica<-rbind(df_ica_opdc,df_ica_ppmi)
df_ica$cohort<-c(rep("ODPC",length(row.names(df_ica_opdc))),rep("PPMI",length(row.names(df_ica_ppmi))))
df_ica$S<-paste0(df_ica$Underly_Variable,sep="_",df_ica$Cat)
df_ica_mean<-correlation_table(df_ica,cat_ppmi_red)
colnames(df_ica_mean)[3]<-"OPDC"
p <- ggplot(df_ica_mean, aes(OPDC, PPMI))+ geom_point(aes(colour = factor(S)),size=10
                                                         ) +geom_text(aes(label=Cat),size=7,hjust=0, vjust=0) + xlim(-0.6, 0.6)+ ylim(-0.6,0.6)+ geom_abline(intercept =0, slope = 1)+theme(axis.text=element_text(size=10),axis.title=element_text(size=18))+ theme(legend.position="none")
p
p_ica<-p
cor.test(df_ica_mean$OPDC,df_ica_mean$PPMI)
```

###  MDS
```{r MDS correlation ,echo=FALSE}
df_mds_odpc<-correlation_lat_var(out$imp,mds_odpc,3,cat_opdc)
df_mds_ppmi<-correlation_lat_var(out2$imp,mds_ppmi,3,cat_ppmi_red)

df_mds<-rbind(df_mds_odpc,df_mds_ppmi)
df_mds$cohort<-c(rep("ODPC",length(row.names(df_mds_odpc))),rep("PPMI",length(row.names(df_mds_ppmi))))
df_mds$S<-paste0(df_mds$Underly_Variable,sep="_",df_mds$Cat)
df_mds_mean<-correlation_table(df_mds,cat_ppmi_red)
colnames(df_mds_mean)[3]<-"OPDC"
p <- ggplot(df_mds_mean, aes(OPDC, PPMI))+ geom_point(aes(colour = factor(S)),size=10
                                                         ) +geom_text(aes(label=Cat),size=7,hjust=0, vjust=0) + xlim(-0.6, 0.6)+ ylim(-0.6,0.6)+ geom_abline(intercept =0, slope = 1)+theme(axis.text=element_text(size=10),axis.title=element_text(size=18))+ theme(legend.position="none")

p
p_mds<-p
cor.test(df_mds_mean$OPDC,df_mds_mean$PPMI)
```

###  PCA
```{r PCA correlation ,echo=FALSE}
df_pca_odpc<-correlation_lat_var(out$imp,pca_odpc,3,cat_opdc)
df_pca_ppmi<-correlation_lat_var(out2$imp,pca_ppmi,3,cat_ppmi_red)

df_pca<-rbind(df_pca_odpc,df_pca_ppmi)
df_pca$cohort<-c(rep("ODPC",length(row.names(df_pca_odpc))),rep("PPMI",length(row.names(df_pca_ppmi))))
df_pca$S<-paste0(df_pca$Underly_Variable,sep="_",df_pca$Cat)
df_pca_mean<-correlation_table(df_pca,cat_ppmi_red)
colnames(df_pca_mean)[3]<-"OPDC"
p <- ggplot(df_pca_mean, aes(OPDC, PPMI))+ geom_point(aes(colour = factor(S)),size=10
                                                         ) +geom_text(aes(label=Cat),size=7,hjust=0, vjust=0) + xlim(-0.6, 0.6)+ ylim(-0.6,0.6)+ geom_abline(intercept =0, slope = 1)+theme(axis.text=element_text(size=10),axis.title=element_text(size=18))+ theme(legend.position="none")
p
p_pca<-p
cor.test(df_pca_mean$OPDC,df_pca_mean$PPMI)
```
