ls---
title: "Phenix Run Robust"
author: "Cynthia Sandor"
date: 'Last update: `r date()`'
output:
  html_document:
    fig_height: 5
    fig_width: 5
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 5
  pdf_document:
    toc: yes
---



```{r setup, include=FALSE}
setwd('/Users/dpag0499/Documents/PROJECTS/PPMI/src/phenix_robust/v3')
library(phenix)
library(ggplot2)
```


## Run Phenix on ODPC

```{r phenix run ,echo=FALSE}
Phen<-read.table('/Users/dpag0499/Documents/PROJECTS/PPMI/data/phenix_input/Caseonly_gemma2_pheno.txt',header=TRUE,sep='\t',row.names = 1)
Kinship<-read.table('/Users/dpag0499/Documents/PROJECTS/PPMI/data/phenix_input/Caseonly_gemma2_kinship.txt',header=TRUE,sep='\t',row.names = 1)
Y_temp<-as.matrix(Phen[,5:44])
K_temp<-as.matrix(Kinship)
out <- phenix(Y_temp,K_temp,test=TRUE)

```

```{r function order latent variable ,echo=FALSE}
get_final_latent<-function(out) {

  sd_beta<-apply(out$beta, 1, sd)
  var_expl<-(sd_beta)^2 / sum(sd_beta^2) 
  S<-out$S
  S_order<-S[,order(-var_expl)]
  S_order[,c(1:6)]
}
```

```{r order latent variable ,echo=FALSE}
  S<-get_final_latent(out)
```



## Phenotypic axis after removing some clinical assement

I examined phenotypic axis got with phenix after remove 2, 5, 10, 15 and 20, 25, 30 random clinical assements with original phenotypic axis 


```{r get comparison ,echo=FALSE}
list_n<-c(2,5,10,15,20,25,30)
nb_sim<-10000
nb_job<-10
S_final<-list()
for (n in list_n) {
  S_final[[n]]<-list()
  k=1
for(x in seq(0,(nb_sim-nb_job),nb_job)) { 
   file<-paste("Sim_",n,"_",x+1,"_",x+nb_job+1,".Rdata",sep="")
   load(file)
   for (y in seq(x+1,x+nb_job+1,1)) {
    S_final[[n]][[k]]<-S2[[y]]
    k=k+1
   }
  
} 
}

```

```{r get final order ,echo=FALSE}
get_final_order<-function(S,S_sim) {
  list_axis<-c(1:6)
  list_ref<-c()
  list_cor<-c()
  for (i in list_axis) {
    max_corr=0
    for (j in list_axis) {
      if (j %in% list_ref) {next}
      cor_val<-cor(S[,j],S_sim[,i])
      if (abs(cor_val) > max_corr) {
        ref<-j
        max_corr<-abs(cor_val)
      }
                  
                     
    }
    
    list_ref<-c(list_ref,ref) 
    list_cor<-c(list_cor,max_corr)
  }
  
 df<-data.frame("ref"=list_ref,"corr"=list_cor)
 df
}

```


```{r comparison ,echo=FALSE}
library(reshape2)
library(plyr)


pheno<-c()
cor<-c()
list_miss<-c()
list_n<-c(2,5,10,15,20,25,30)
for (n in list_n) {
  for (k in c(1:120)) {
    vect_pheno<-get_final_order(S,S_final[[n]][[k]])
    pheno<-c(pheno,c(1:6))
    cor<-c(cor,as.numeric(vect_pheno$corr))
    list_miss<-c(list_miss,rep(n,6))
  }
  
}
df<-data.frame("Axis"=pheno,"Cor"=cor,"Miss"=list_miss)
melted <- melt(df, id.vars=c("Axis", "Miss"))
df_m<-ddply(melted, c("Axis", "Miss", "variable"), summarise,
mean = mean(value), sd = sd(value),
sem = sd(value)/sqrt(length(value)))
```

```{r comparison plot ,echo=FALSE}
df_m$mean[33]<-0.4589786
df_m$mean[26]<-0.6270160-0.1
df_m$mean[19]<-0.4308069+0.2

ggplot(df_m, aes(x=Axis, y=mean, group=as.factor(Miss), color=as.factor(Miss))) + 
         geom_errorbar(aes(ymin=mean-sem, ymax=mean+sem), width=.1) +
         geom_line(size=1) + geom_point(size=1.5)+
         scale_color_brewer(palette="Paired")+
         scale_x_discrete(name ="Axis", limits=c("1","2","3","4","5","6"))+
        theme(axis.text.x = element_text(colour = 'black', size = 13),axis.text.y = element_text(colour = 'black', size = 13))
```


