---
title: "Comparison PPMI ODPC multiple methods add ICA "
author: "Cynthia Sandor"
date: 'Last update: `r date()`'
output:
  html_document:
    fig_height: 5
    fig_width: 5
    theme: cerulean
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
      toc_depth: 5
  pdf_document:
    toc: yes
---


```{r setup, include=FALSE}
setwd('/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_multiple_approaches')
library(phenix)
library(ggplot2)
library(gplots)
library(reshape)
load("/Users/dpag0499/Documents/PROJECTS/PPMI/data/Rdata/ppmi.Rdata")
load("/Users/dpag0499/Documents/PROJECTS/PPMI/src/comparison_ppmi_odpc_phenix/ODPC.Rdata")
```

```{r get category, include=FALSE}
cat_opdc<-read.table("category_opdc_red.txt",h=T,sep="\t")
cat_ppmi<-read.table("category_ppmi.txt",h=T,sep="\t")
```



```{r functions, include=FALSE}



correlation_lat_var<-function(Y,S,n_axes) {
  n_clinical<-length(colnames(Y))
  list_axes<-c(1:n_axes)
  mm <- matrix(NA,n_clinical,n_axes)

  for (i in 1:n_axes) {
    for (j in 1:n_clinical) {
      t<-cor.test(Y[,j],S[,i])
      mm[j,i]<-as.numeric(t$estimate)
    }
  }
    df_mm<-as.data.frame(mm)
    colnames(df_mm)<-paste0("V",c(1:n_axes))
    row.names(df_mm)<-colnames(Y)
    df_mm$Clinical<-row.names(df_mm)
    
    df_mm
}

reshape_correlation<-function(df,cat){
  df_mm<-merge(df,cat[,-c(2)],by=c("Clinical"))
  df_mm_s<-melt(df_mm,id=c("Clinical","Cat","Sgn"))
  df_mm_s$value2<-df_mm_s$value*df_mm_s$Sgn
  df_mm_s
}

reshape_correlation<-function(df,cat){
  df_mm<-merge(df,cat[,-c(2)],by=c("Clinical"))
  df_mm_s<-melt(df_mm,id=c("Clinical","Cat","Sgn"))
  df_mm_s$value2<-df_mm_s$value*df_mm_s$Sgn
  df_mm_s
}

```





## Compute MDS/PCA/ICA/phenix


### Phenix
```{r phenix ,echo=FALSE}
phenix_odpc<-out$S
phenix_ppmi<-out3$S
phenix_ppmi[,2]<--phenix_ppmi[,2]
phenix_ppmi[,3]<-phenix_ppmi[,4]
```



## Compute correlation underlying variable & phenotypic axes

As PPMI and ODPC do not use the same clinical assessments, I classified into different categories:
(1) Motors
(2) Cognitive
(3) Anxiety-Deepression
(4) Autonomic
(5) Sleep
(6) Olfactive


These categories have been defined by PPMI cnsortium.

(1) I computed the correlation (pearson) between clinical assements with principal variable for different categories of phenotype in both cohort. 
(2) I computed the mean of correlation by category of phenotype
(3) I compared these correlation by category p




###  Phenix

```{r Phenix correlation ,echo=FALSE}
## correlation
df_phenix_odpc<-correlation_lat_var(out$imp,phenix_odpc,3)
df_phenix_ppmi<-correlation_lat_var(out3$imp,phenix_ppmi,3)

## reshape and add category
df_phenix_odpc_r<-reshape_correlation(df_phenix_odpc,cat_opdc)
df_phenix_ppmi_r<-reshape_correlation(df_phenix_ppmi,cat_ppmi)

## compute mean
df_phenix_odpc_mean<-ddply(df_phenix_odpc_r,~Cat+variable,summarise,mean=mean(value2))
df_phenix_ppmi_mean<-ddply(df_phenix_ppmi_r,~Cat+variable,summarise,mean=mean(value2))

## final table
df_phenix<-change_sgn_unique_table(df_phenix_odpc_mean,df_phenix_ppmi_mean)

```








###  All Together
```{r all analysis together correlation ,echo=FALSE}
df_analysis<-df_phenix
df_analysis$Method<-rep("Phenix",dim(df_phenix)[1])
df_analysis_temp<-df_ica
df_analysis_temp$Method<-rep("ICA",dim(df_ica)[1])
df_analysis<-rbind(df_analysis,df_analysis_temp)
df_analysis_temp<-df_mds
df_analysis_temp$Method<-rep("MDS",dim(df_mds)[1])
df_analysis<-rbind(df_analysis,df_analysis_temp)
df_analysis_temp<-df_pca
df_analysis_temp$Method<-rep("PCA",dim(df_pca)[1])
df_analysis<-rbind(df_analysis,df_analysis_temp)


```


